### 2016-02-08 Manuscript/BiologyLetters/CoverLetter.tex
re committed files

Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: % NC: What is the title?

@@

The Total Evidence method allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
%TG: it changes from here:
This method is extremely promising because it allows us to use all the available data. 
However, to efficiently apply this method, it is crucial to have enough morphological data for living species to allow the fossils to branch correctly in the phylogeny (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution).

>>>>

Please find attached our research article, ``Assessment of cladistic data availability for living mammals'', for consideration in the Biology Letters Special Issue: ``Putting fossils in trees''.

@@

The Total Evidence method allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al. 2012 Syst Biol).
This method is extremely promising because it allows us to use all the available data. 
However, to efficiently apply this method, it is crucial to have enough morphological data for living species to allow the fossils to branch correctly in the phylogeny (Guillerme and Cooper 2016, Molecular Phylogenetics and Evolution).

### 2016-02-08 Manuscript/BiologyLetters/Missing_mammals-draft.tex
re committed files

It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e. we did specifially not include any matrices that were only available in paper format (e.g. printed in books), non-reusable format (e.g. an image of the matrix) %TG: yes... There were quiet a few of them!
or/and matrices available only upon request (e.g. by emailing the authors).
Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.

Despite the absence of good cladistic data coverage for living mammals, %TG: I think leaving cladistic here is OK no?
 the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses. 
Following the recommendations in \cite{GuillermeCooper}, we need to code anatomical characters for as many living species possible. 
Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.

>>>>

It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e. we did specifically not include any matrices that were only available in paper format (e.g. printed in books), non-reusable format (e.g. an image of the matrix) %TG: yes... There were quiet a few of them!
or/and matrices available only upon request (e.g. by emailing the authors).
Such matrices are likely to not be included in Total Evidence analysis due to the difficulty of obtaining them.

Despite the absence of good cladistic data coverage for living mammals, %TG: I think leaving cladistic here is OK no?
 the Total Evidence method still seems to be the most promising way of combining living and fossil data for macroevolutionary analyses.
It is also encouraging to note that the sampling effort is consistent with diversity at each taxonomic level and there is therefore no direct bias in sampling at any taxonomic level (ESM 2, Figure 1).
Following the recommendations in \cite{GuillermeCooper}, we need to code anatomical characters for as many living species possible. 
Fortunately, data for living mammals is usually readily available in natural history collections, therefore, we propose that an increased effort be put into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
Such an effort would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.

### 2016-02-08 Manuscript/Missing_mammals-draft.tex
re committed files

\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$$^,$$^3$}\\
\noindent {\small \it 
$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\ %Should we remove that one for word count?
$^3$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
\end{center}
\medskip
\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\  
\vspace{1in}

%Line numbering
\modulolinenumbers[1]
\linenumbers

>>>>

\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^*$ and Natalie Cooper$^1$$^,$$^2$}\\
\noindent {\small \it 
$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
$^2$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.}\\
\end{center}
\medskip
\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\  
\vspace{1in}

%Line numbering REMOVE FOR WORDCOUNT
%\modulolinenumbers[1]
%\linenumbers

### 2016-02-08 Manuscript/Supplementary/Results_1c/Results_1cTable_results.tex
re committed files

\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels. The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. *p \textless 0.05; **p \textless 0.01; ***p \textless 0.001.} \\ 

>>>>

\caption{Number of taxa with available cladistic data for mammalian orders at three taxonomic levels (without any character threshold; results from the 286 matrices). The coverage represents the proportion of taxa with available morphological data. The left vertical bar represents 25\% of available data (``low'' coverage if \textless 25\%); The right vertical bar represents 75\% of available data (``high'' coverage if \textgreater 75\%). When the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) are negative, taxa are more phylogenetically dispersed than expected by chance; when NRI or NTI are positive, taxa are more phylogenetically clustered by expected by chance. Significant NRI or NTI are highlighted in bold. One star (*) represents a p-value between 0.05 and 0.005; two starts between 0.005 and 0.0005 and three stars a p-value less than 0.0005.} \\ 

### 2016-02-08 Manuscript/proto-draft.tex
re committed files

\subsection{Mammalian orders terms}
The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt. The file containing the meta names is the latin name files without the following suffixes: Mammal     @ia
Monothrem     @ata
Marsupial     @ia
Placental     @ia
Macroscelid     @ea
Afrosoricid     @a
Tubulident     @ata
Hyracoid     @ea
Proboscid     @ea
Siren     @ia
Pilos     @a
Cingul     @ata
Scandent     @ia
Dermopter     @a
Primat     @es
Lagomorph     @a
Rodent     @ia
Erinaceomorph     @a
Soricomorph     @a
Cetac     @ea
Artiodactyl     @a
Chiropter     @a
Perissodactyl     @a
Pholidot     @a
Carnivor     @a
Didelphimorph     @ia
Paucitubercul     @ata
Microbiother     @ia
Dasyuromorph     @ia
Peramelemorph     @ia
Notoryctemorph     @ia
Diprotodont     @ia
Monotrem     @ata

>>>>

\subsection{Mammalian orders terms}
The searched mammalian order terms are available in search_terms_latin.txt or search_terms_meta.txt. The file containing the meta names is the Latin name files but with replacing the latin suffixes ([ia|ata|ea|a]) by a joker character (*) and by replacing the first letter by a upper/lower case meta character (e.g. [Aa]).

\subsubsection{Ross Mounce data set}
I selected all the matrices containing at least one of the mammalian orders names from Ross Mounce GitHub cladistic-data/nexus_files repository (accessed on the 02/12/2014).
